---
title: "Case Fatality Graphs"
author: "Anna"
date: "01/04/2021"
output:   
  html_document:
    keep_md: true
---

These graphs show case fatality rates by sex, imd & age-group

```{r setup, include=FALSE, warning=FALSE}

library(ggplot2)
library(knitr)
library(Rcpp)
library(fst)
library(data.table)
if (!require("PHEindicatormethods")) install.packages("PHEindicatormethods")
library(PHEindicatormethods)
library(RColorBrewer)
library(ggthemes)
library(scales)

#DSR package now archived
#packageurl <- "https://cran.r-project.org/src/contrib/Archive/dsr/dsr_0.2.2.tar.gz"
#install.packages(packageurl, repos=NULL, type="source")
library(dsr)


data.table::setDTthreads(10) #this is so that don't use all the processors 
```

# colour palettes etc
```{r include=FALSE}
cbPalette <- c("#999999", "#E69F00", "#56B4E9", "#009E73", "#F0E442", "#0072B2", "#D55E00", "#CC79A7")

cbPaletteMod <- c("#999999", "#E69F00", "#56B4E9", "#009E73", "#D55E00", "#CC79A7")



mypalette <-brewer.pal(7,"GnBu")
mypalette #don't want the faintest 2 
mypalette <- c("#A8DDB5" ,"#7BCCC4" ,"#4EB3D3" ,"#2B8CBE", "#08589E")

xaxisyear <- c(2004, 2007, 2010, 2013, 2016, 2019)



mypalette2 <-brewer.pal(9,"GnBu")
mypalette2 #don't want the faintest 2 
mypalette2 <- c("#CCEBC5", "#A8DDB5" ,"#7BCCC4" ,"#4EB3D3" ,"#2B8CBE", "#0868AC" , "#084081")

```

# load the data
```{r echo=FALSE, out.width = '100%'}
data_dir_CPRD <-
  function(x = character(0))
    paste0("/mnt/", Sys.info()[["user"]],
      "/UoL/CPRD2019mm/Data May 2020/", x)

data_dir_lookup <-
  function(x = character(0))
    paste0("/mnt/", Sys.info()[["user"]],
           "/UoL/CPRD2019mm/Dictionaries etc/", x)


combi_mm <- read_fst(data_dir_CPRD("combi_mm_detailed.fst"), as.data.table = T)
combi_mm[, `:=` (
  imd = factor(imd, levels = c(1,2,3,4,5)), 
  bmm = factor(bmm),
  cmm = factor(cmm),
  cmm3plus = factor(cmm3plus))]



```

# Table for casefatality
```{r}
cf_tab <- rbind(combi_mm[bmm != 0 & bmm_days >0, 
                            .(N = sum(censorreason == 1 ), 
                              days_atrisk = sum(bmm_days), 
                              outcome = (sum(censorreason == 1)/ (sum(bmm_days)/365.24))*10000), 
                            keyby = c("year", "gender", "agegrp5", "agegrp10_simple","imd", "region")][
                              , type := "Basic"], 
               combi_mm[cmm != 0 & cmm_days >0, 
                            .(N = sum(censorreason == 1 ), 
                              days_atrisk = sum(cmm_days), 
                              outcome = (sum(censorreason == 1)/ (sum(cmm_days)/365.24))*10000),
                            keyby = c("year", "gender", "agegrp5", "agegrp10_simple","imd", "region")][
                              ,   type := "Complex"],
               combi_mm[cmm3plus != 0 & cmm3plus_days >0, 
                            .(N = sum(censorreason == 1 ), 
                              days_atrisk = sum(cmm3plus_days), 
                              outcome = (sum(censorreason == 1)/ (sum(cmm3plus_days)/365.24))*10000),
                            keyby = c("year", "gender", "agegrp5", "agegrp10_simple","imd", "region")][
                              ,   type := "Complex 3+"])


```


#### Looking at case fatality 
Has declined over the period

```{r echo=FALSE, out.width = '100%'}

ggplot(
    cf_tab[, .(outcome = sum(N) / (sum(days_atrisk)/ 365.24) *10000),
           by = c("year",  "type" )], 
    aes(year, outcome, col = type) ) +
    geom_smooth(size = 0.5, se = FALSE) +
    expand_limits(y=0) +    
    scale_x_continuous(name="Year", limits=c(2003.5, 2020), breaks = xaxisyear) + 
    scale_y_continuous(name="Crude Case Fatality Rate\n (per 10,000 person-years)", 
                       labels = comma_format()) +
    theme_few() +theme(axis.text.x = element_text(size = 8),
                       legend.title = element_text( size=10)) +
    labs(colour = "MM Type")  



cf_tab[year %in% c(2004, 2019), 
       .(outcome = sum(N) / (sum(days_atrisk)/ 365.24) *10000),
       by = c("year",  "type" )]

```


# By sex
Has declined over the period 
```{r echo=FALSE, out.width = '100%'}
ggplot(
    cf_tab[gender != "Indeterminate", .(outcome = sum(N) / (sum(days_atrisk)/ 365.24) *10000),
           by = c("year",  "type", "gender" )], 
    aes(year, outcome, col = type) ) +
    geom_smooth(size = 0.5, se = FALSE) +
  facet_grid(cols = vars(gender)) +
    expand_limits(y=0) +    
    scale_x_continuous(name="Year", limits=c(2003.5, 2020), breaks = xaxisyear) + 
    scale_y_continuous(name="Crude Case Fatality Rate\n (per 10,000 person-years)", 
                       labels = comma_format()) +
    theme_few() +theme(axis.text.x = element_text(size = 8),
                       legend.title = element_text( size=10)) +
    labs(colour = "MM Type")  


cf_tab[gender != "Indeterminate", .(outcome = sum(N) / (sum(days_atrisk)/ 365.24) *10000),
           keyby = c("year",  "type" )]

```


# By imd
Has declined over the period - some narrowing of inequalities??? Although could this be be to do with age...
```{r echo=FALSE, out.width = '100%'}
ggplot(
    cf_tab[imd != "", .(outcome = sum(N) / (sum(days_atrisk)/ 365.24) *10000),
           by = c("year",  "type", "imd" )], 
    aes(year, outcome, col = imd) ) +
    geom_smooth(size = 0.5, se = FALSE) +
  facet_grid(cols = vars(type)) +
    expand_limits(y=0) +    
    scale_x_continuous(name="Year", limits=c(2003.5, 2020), breaks = xaxisyear) +
    scale_y_continuous(name="Crude Case Fatality Rate\n (per 10,000 person-years)", 
                       labels = comma_format()) +
  scale_colour_manual(values = mypalette) +
    theme_few() +theme(axis.text.x = element_text(size = 8),
                       legend.title = element_text( size=10)) +
    labs(colour = "IMD") 

```


# By 5year age croup 
5year agegroup cf rates haven't really declined... 
```{r echo=FALSE, out.width = '100%'}
 ggplot(
    cf_tab[imd != "", .(outcome = sum(N) / (sum(days_atrisk)/ 365.24) *10000),
           by = c("year",  "type", "agegrp5" )], 
    aes(year, outcome, col = agegrp5) ) +
    geom_smooth(size = 0.5, se = FALSE) +
  facet_grid(cols = vars(type)) +
    expand_limits(y=0) +    
    scale_x_continuous(name="Year", limits=c(2003.5, 2020), breaks = xaxisyear) +
    scale_y_continuous(name="Crude Case Fatality Rate\n (per 10,000 person-years)", 
                       labels = comma_format()) +
    theme_few() +theme(axis.text.x = element_text(size = 8),
                       legend.title = element_text( size=10)) +
    labs(colour = "5-year age-group")       

```

# By 10year age croup 
```{r echo=FALSE, out.width = '100%'}
 ggplot(
    cf_tab[imd != "", .(outcome = sum(N) / (sum(days_atrisk)/ 365.24) *10000),
           by = c("year",  "type", "agegrp10_simple" )], 
    aes(year, outcome, col = agegrp10_simple) ) +
    geom_smooth(size = 0.5, se = FALSE) +
  facet_grid(cols = vars(type)) +
    expand_limits(y=0) +    
    scale_x_continuous(name="Year", limits=c(2003.5, 2020), breaks = xaxisyear) +
    scale_y_continuous(name="Crude Case Fatality Rate\n (per 10,000 person-years)", 
                       labels = comma_format()) +
    theme_few() +theme(axis.text.x = element_text(size = 8),
                       legend.title = element_text( size=10)) +
    labs(colour = "10-year age-group")    
```

```{r}
ggplot(
    cf_tab[gender != "Indeterminate" & type != "Complex 3+",
           .(outcome = sum(N) / (sum(days_atrisk)/ 365.24) *10000), by = c("year", "agegrp10_simple", "type", "gender")], 
    aes(year, outcome, col = agegrp10_simple) ) +
    geom_smooth(size = 0.5, se = FALSE) +
    facet_grid(type ~ gender) +  
    expand_limits(y=0) +    
    scale_x_continuous(name="Year", limits=c(2003.5, 2020), breaks = xaxisyear) + 
    scale_y_continuous(name="Crude Case Fatality Rate\n (per 10,000 person-years)", 
                       labels = comma_format()) +
    theme_few() +theme(axis.text.x = element_text(size = 8),
                       legend.title = element_text( size=10)) +
    #guides(colour = "none") +
    scale_colour_manual(values = mypalette2)  +
    #scale_colour_brewer(type = 'qual', palette = "Dark2")  +
    labs(colour = "10-year\n age-group")  +
  guides(colour = guide_legend(reverse=TRUE))
#ggsave("prelim_graphs/cfr_sex_age.svg") 

ggplot(
    cf_tab[gender != "Indeterminate",
           .(outcome = sum(N) / (sum(days_atrisk)/ 365.24) *10000), by = c("year", "agegrp10_simple", "type", "gender")], 
    aes(year, outcome, col = agegrp10_simple) ) +
    geom_smooth(size = 0.5, se = FALSE) +
    facet_grid(gender ~ type ) +  
    expand_limits(y=0) +    
    scale_x_continuous(name="Year", limits=c(2003.5, 2020), breaks = xaxisyear) + 
    scale_y_continuous(name="Crude Case Fatality Rate\n (per 10,000 person-years)", 
                       labels = comma_format()) +
    theme_few() +theme(axis.text.x = element_text(size = 8),
                       legend.title = element_text( size=10)) +
    #guides(colour = "none") +
    scale_colour_manual(values = mypalette2)  +
    #scale_colour_brewer(type = 'qual', palette = "Dark2")  +
    labs(colour = "10-year\n age-group")  +
  guides(colour = guide_legend(reverse=TRUE))
#ggsave("prelim_graphs/prelim_sa/cfr_sex_age_3p.svg") 
```

# Age-standardised

```{r}
eustandpop <- fread(file = data_dir_lookup("eustandardpop5yrband.csv"), 
                    header = TRUE, sep = ",") #European Standardised pop 
eustandpop[, gender := as.factor(ifelse(gender== "Male", "Men", "Women"))] 
setnames(eustandpop, c("stand.pop", "age.group"), c("pop", "agegrp5"))



dsr_a_imd <- function(x, y){
  DT <- cf_tab[imd == x & gender != "Indeterminate" & type == y, 
                                .(N = sum(N),
                atrisk = sum(days_atrisk)/365.24),
               by = c("year", "gender", "agegrp5")]
  z <-  rbind(
  as.data.table( dsr(
    data= DT[gender == "Men", ],
    event=N,
    fu=atrisk,
    subgroup=year, age.group, type,
    refdata=eustandpop[gender == "Men",],
    method="gamma",
    sig=0.95,
    mp=10000,
    decimals=4))[, gender := "Men"], 
  as.data.table( dsr(
        data= DT[gender == "Women", ],
        event=N,
        fu=atrisk,
        subgroup=year, age.group, type,
        refdata=eustandpop[gender == "Women",],
        method="gamma",
        sig=0.95,
        mp=10000,
        decimals=4))[, gender := "Women"] )
  z[,`:=` (imd = x)]
  z
}

bmm_cfr_ast_imd <- rbind(
  dsr_a_imd(1, "Basic"),
  dsr_a_imd(2, "Basic"),
  dsr_a_imd(3, "Basic"),
  dsr_a_imd(4, "Basic"),
  dsr_a_imd(5, "Basic")
)


cmm_cfr_ast_imd <- rbind(
  dsr_a_imd(1, "Complex"),
  dsr_a_imd(2, "Complex"),
  dsr_a_imd(3, "Complex"),
  dsr_a_imd(4, "Complex"),
  dsr_a_imd(5, "Complex")
)


cmm3plus_cfr_ast_imd <- rbind(
  dsr_a_imd(1, "Complex"),
  dsr_a_imd(2, "Complex"),
  dsr_a_imd(3, "Complex"),
  dsr_a_imd(4, "Complex"),
  dsr_a_imd(5, "Complex")
)
rm(dsr_a_imd)


combi_cfr_ast_imd <-rbind(bmm_cfr_ast_imd[, type := "Basic"], 
                          cmm_cfr_ast_imd[, type := "Complex"], 
                          cmm3plus_cfr_ast_imd[, type := "Complex 3+"])

setnames(combi_cfr_ast_imd, 
         c("Subgroup", "Std Rate (per 10000)", "95% LCL (Std)", "95% UCL (Std)"), 
         c("year", "standoutcome", "standlower", "standupper"))
combi_cfr_ast_imd[, imd := factor(imd)]


ggplot(combi_cfr_ast_imd[type != "Complex 3+"]) + theme_few() +
  geom_smooth(
    mapping = aes(x = year, y = standoutcome, colour = imd),
    se = FALSE,
    size = 0.5
  ) +
  facet_grid(type ~ gender) +
  ylab("Age-standardised case fatality rate\n (per 10,000 person-years)") +
  expand_limits(y = 0) +
  scale_x_continuous(name = "Year",
                     limits = c(2003.5, 2019.5),
                     breaks = xaxisyear) +
  theme(
    axis.text.x = element_text(size = 8),
    legend.background = element_rect(
      fill = '#FBFBFB',
      size = 0.5,
      linetype = "solid"
    ),
    legend.title = element_text(size = 10)
  ) +
  scale_colour_manual(values = mypalette)  +
  labs(colour = "IMD")  +
  guides(colour = guide_legend(reverse = TRUE))

#ggsave("prelim_graphs/cfr_combi_asst_sex_imd.svg") 

ggplot(combi_cfr_ast_imd) + theme_few() +
  geom_smooth(
    mapping = aes(x = year, y = standoutcome, colour = imd),
    se = FALSE,
    size = 0.5
  ) +
  facet_grid(gender ~ type) +
  ylab("Age-standardised case fatality rate\n (per 10,000 person-years)") +
  expand_limits(y = 0) +
  scale_x_continuous(name = "Year",
                     limits = c(2003.5, 2019.5),
                     breaks = xaxisyear) +
  theme(
    axis.text.x = element_text(size = 8),
    legend.background = element_rect(
      fill = '#FBFBFB',
      size = 0.5,
      linetype = "solid"
    ),
    legend.title = element_text(size = 10)
  ) +
  scale_colour_manual(values = mypalette)  +
  labs(colour = "IMD")   +
  guides(colour = guide_legend(reverse = TRUE))
#ggsave("prelim_graphs/prelim_sa/cfr_combi_asst_sex_imd.svg") 

combi_cfr_ast_imd
```

