---
title: "Descriptive Graphs"
author: "Anna"
date: "10/16/2020"
output: 
  html_document:
    keep_md: true
---

```{r setup, include=FALSE}
library(ggplot2)
library(knitr)
library(fst)
library(data.table)
library(tidyverse)
library(ggthemes)


data.table::setDTthreads(10) #this is so that don't use all the processors 

xaxisyear <- c(2004, 2008, 2012, 2016, 2020)

```

# load the data
```{r echo=FALSE, out.width = '100%'}
data_dir_CPRD <-
  function(x = character(0))
    paste0("/mnt/", Sys.info()[["user"]],
      "/UoL/CPRD2019mm/Data May 2020/", x)

combi_mm <- read_fst(data_dir_CPRD("combi_mm_detailed.fst"), as.data.table = T)
combi_mm[, `:=` (
  imd = factor(imd), 
  bmm = factor(bmm),
  cmm = factor(cmm))]


```





# Descriptive

```{r}

require(scales)
ggplot(combi_mm[, .N , keyby = c("year", "gender") ], 
                     aes(x =year, y = N, fill = gender)) + 
geom_bar(stat= "identity", position=position_stack()) + 
  scale_y_continuous(labels = comma)
#ggsave("prelim_graphs/descr_num.png") 

ggplot(combi_mm[, .N , keyby = c("year", "gender", "agegrp10_simple") ], 
       aes(x =year, y = N, fill = agegrp10_simple)) +
  geom_bar(stat= "identity", position=position_fill()) + 
  facet_grid(cols = vars(gender)) +   
  ylab("proportion") 
#ggsave("prelim_graphs/descr_sex_age.png") 

ggplot(combi_mm[imd != "", .N , keyby = c("year", "imd") ], 
       aes(x =year, y = N, fill = imd)) + 
  geom_bar(stat= "identity", position=position_fill()) +   
  ylab("proportion")

ggplot(combi_mm[imd != "", .N , keyby = c("year", "imd", "gender") ], 
       aes(x =year, y = N, fill = imd)) + 
  geom_bar(stat= "identity", position=position_fill()) +   
    facet_grid(cols = vars(gender)) +   
  ylab("proportion")
#ggsave("prelim_graphs/descr_sex_imd.png") 

ggplot(combi_mm[imd != "", .N , keyby = c("year", "imd", "agegrp10_simple") ], 
       aes(x =year, y = N, fill = agegrp10_simple)) +
  geom_bar(stat= "identity", position=position_fill()) + 
  facet_grid(cols = vars(imd)) +   
  ylab("proportion") +   
  theme_few() +
  theme(axis.text.x = element_text(size = 6, angle = 90, hjust = 1)) +
  labs(fill = "Age group") 
ggplot(combi_mm[imd != "", .N , keyby = c("year", "imd", "agegrp10_simple") ], 
       aes(x =year, y = N, fill = imd)) +
    geom_bar(stat= "identity", position=position_fill()) + 
    facet_grid(cols = vars(agegrp10_simple)) +   
    ylab("proportion") +   
    theme_few() +
    theme(axis.text.x = element_text(size = 6, angle = 90, hjust = 1)) +
    labs(fill = "Age group") 
#ggsave("prelim_graphs/descr_imd_age.png") 

```


# Indeterminate Sex 
```{r}
ggplot(combi_mm[gender == "Indeterminate", 
                .N , 
                keyby = c("year", "agegrp10_simple")],
       aes(x = year, y = N, fill = agegrp10_simple)) +
  geom_bar(stat = "identity", position = position_stack()) +
  scale_y_continuous(name = "Number individuals registered", 
                     limits = c(0, 200))
#ggsave("prelim_graphs/descr_sexI_age.png")

ggplot(combi_mm[gender == "Indeterminate" & imd != "", 
                .N , 
                keyby = c("year", "imd") ], 
       aes(x =year, y = N, fill = imd)) + 
  geom_bar(stat= "identity", position=position_fill()) +   
  ylab("proportion") #+ 
  #scale_y_continuous(name="Number individuals registered", limits = c(0, 200))
#ggsave("prelim_graphs/descr_sexI_imd.png") 

```




# IMD Missing
```{r}
ggplot(combi_mm[gender != "Indeterminate" &
                  imd == "", .N , keyby = c("year", "gender")],
       aes(x = year, y = N, fill = gender)) +
  geom_bar(stat = "identity", position = position_stack()) +
  scale_y_continuous(name = "Number individuals registered", limits = c(0, 750))
#ggsave("prelim_graphs/noimd_sex.png")


ggplot(combi_mm[imd == "", .N , keyby = c("year", "agegrp10_simple") ], 
       aes(x =year, y = N, fill = agegrp10_simple)) + 
  geom_bar(stat= "identity", position=position_fill()) +   
  ylab("proportion") #+ 
  #scale_y_continuous(name="Number individuals registered", limits = c(0, 200))
#ggsave("prelim_graphs/noimd_age.png") 

```

#Looking at nmbers by sex
```{r}

combi_mm[, table(year, gender)]
combi_mm[, as.list(round(table(gender)/.N*100, 2)), keyby = .(year)]

combi_mm[year %in% c(2004, 2019), 
         table(year, gender)]
combi_mm[year %in% c(2004, 2019), 
         as.list(round(table(gender)/.N*100, 2)), keyby = .(year)]
```

#Looking at nmbers in different agegroups 
```{r}
combi_mm[year == 2004, summary(age)]
combi_mm[year == 2019, summary(age)]

combi_mm[, table(year, agegrp10_simple)]
combi_mm[, as.list(round(table(agegrp10_simple)/.N*100, 2)), keyby = .(year)]

combi_mm[year %in% c(2004, 2019), 
         table(year, agegrp10_simple)]
combi_mm[year %in% c(2004, 2019), 
         as.list(round(table(agegrp10_simple)/.N*100, 2)), keyby = .(year)]


ggplot(combi_mm[imd != "", .N , keyby = c("year", "imd", "agegrp10_simple") ], 
       aes(x =year, y = N, fill = agegrp10_simple)) +
    geom_bar(stat= "identity", position=position_stack()) + 
    facet_grid(cols = vars(imd)) +   
    ylab("proportion") +   
    theme(axis.text.x = element_text(size = 6, angle = 90, hjust = 1))



ggplot(combi_mm[imd != "", .N , keyby = c("year", "imd", "agegrp10_simple") ], 
       aes(x =year, y = N, fill = agegrp10_simple)) +
    geom_bar(stat= "identity", position=position_stack()) + 
    facet_grid(cols = vars(imd)) +   
    ylab("proportion") +   
  theme_few() +
    theme(axis.text.x = element_text(size = 6, angle = 90, hjust = 1))


ggplot(combi_mm[imd != "", .N , keyby = c("year", "imd", "agegrp10_simple") ], 
       aes(x =year, y = N, fill = imd)) +
    geom_bar(stat= "identity", position=position_stack()) + 
    facet_grid(cols = vars(agegrp10_simple)) +   
    ylab("proportion") +   
  theme_few() +
    theme(axis.text.x = element_text(size = 6, angle = 90, hjust = 1))




ggplot(combi_mm[agegrp10_simple == "18-29" &
                  imd != "", .N , keyby = c("year", "imd")],
       aes(x = year, y = N, fill = imd)) +
  geom_bar(stat = "identity", position = position_stack()) +
  ylab("proportion") +
  theme_few() +
  theme(axis.text.x = element_text(
    size = 6,
    angle = 90,
    hjust = 1
  ))


```

#Looking at nmbers by imd
```{r}

combi_mm[, table(year, imd)]
combi_mm[, as.list(round(table(imd)/.N*100, 2)), keyby = .(year)]

combi_mm[year %in% c(2004, 2019), 
         table(year, imd)]
combi_mm[year %in% c(2004, 2019), 
         as.list(round(table(imd)/.N*100, 2)), keyby = .(year)]
```


#Looking at nmbers by region
```{r}

combi_mm[, table(year, region)]
combi_mm[, as.list(round(table(region)/.N*100, 2)), keyby = .(year)]

combi_mm[year %in% c(2004, 2019), 
         table(year, region)]
combi_mm[year %in% c(2004, 2019), 
         as.list(round(table(region)/.N*100, 2)), keyby = .(year)]
```

#Looking at bmm & cmm totals by year
```{r}

combi_mm[, .(N=sum(bmm != 0), pc = sum(bmm != 0)/.N), keyby = year]
combi_mm[, .(N=sum(cmm != 0), pc = sum(cmm != 0)/.N), keyby = year]

```

# looking at time in study
```{r}
study_pats <- read_fst(data_dir_CPRD("study_pats.fst"), as.data.table = TRUE)

study_pats[, enterstudy := as.IDate(ifelse(
  reg1yr < 2004-01-01, #setting start date to 1Jan2004 as earliest 
  as.IDate((paste0("01012004")), format = "%d%m%Y"), 
  reg1yr ), format = "Y%/m%/d%")]
study_pats[, enterstudy := as.IDate(ifelse(
  reg1yr < 2004-01-01, #setting start date to 1Jan2004 as earliest 
  as.IDate((paste0("01012004")), format = "%d%m%Y"), 
  reg1yr ), format = "Y%/m%/d%")]


study_pats[censordate - enterstudy >=0, 
        summary((censordate - enterstudy)/365.24)]

```




# Looking at start vs end status - need to add back in censordate
```{r}
setkey(combi_mm, patid, year)
options(digits = 3)
first_year <- combi_mm[combi_mm[, .I[1], keyby = c("patid")]$V1]
combi_mm[study_pats, on = 'patid', `:=`
         (censordate = i.censordate, censorreason = i.censorreason)]
last_year <-
  combi_mm[year(censordate) == year | #want the final year of reg
             (year == "2019" & year(censordate) == 2020),]
first_year[last_year, on = 'patid', `:=` (end_bmm = i.bmm, end_cmm = i.cmm)]

first_year[, .(
  h_h = sum(end_bmm == 0) / sum(bmm != 2) * 100,
  #Start healthy (includes incidentbmm in 1st year) stay healthy
  h_bmm = sum(end_bmm != 0 &
                bmm != 2 & end_cmm == 0) / sum(bmm != 2) * 100,
  #healthy --> bmm
  h_cmm = sum(end_cmm != 0 &
                bmm != 2) / sum(bmm != 2) * 100,
  #healthy --> cmm
  bmm_cmm = sum(end_cmm != 0 &
                  bmm == 2 & cmm == 0) / sum(bmm == 2) * 100
)] #bmm --> cmm


#Doing in 50+ only for comparison with Strauss paper
combi_mm_50plus <- combi_mm[age>49,]
first_year_50plus <- combi_mm_50plus[combi_mm_50plus[, .I[1], 
                                                     keyby = c("patid")]$V1]
first_year_50plus[last_year, on = 'patid', `:=` 
                  (end_bmm = i.bmm, end_cmm = i.cmm)]

first_year_50plus[, #Start healthy (incl incidentbmm in 1st year) stay healthy: 
                  .(h_h = sum(end_bmm == 0)/sum(bmm != 2)*100, 
               h_bmm = sum(end_bmm != 0 & bmm != 2 & end_cmm == 0)/
                 sum(bmm != 2)*100, #healthy --> bmm
               h_cmm = sum(end_cmm != 0 & bmm != 2)/
                 sum(bmm != 2)*100, #healthy --> cmm
               bmm_cmm = sum(end_cmm != 0 & bmm == 2 & cmm == 0 )/
                 sum(bmm == 2)*100)] #bmm --> cmm


```


# Mean/sd/median/iqr # of conditions in those with basic/complex multimorbidity 
```{r}
options(digits = 2)


combi_mm[bmm != 0, .(mean = mean(n_cond),
                    sd = sd(n_cond), 
                    lower = quantile(n_cond, .25, na.rm=TRUE),
                  middle = quantile(n_cond, .50, na.rm=TRUE),
                  upper = quantile(n_cond, .75, na.rm=TRUE)), keyby = year]
combi_mm[cmm != 0, .(mean = mean(n_cond),
                    sd = sd(n_cond), 
                    lower = quantile(n_cond, .25, na.rm=TRUE),
                  middle = quantile(n_cond, .50, na.rm=TRUE),
                  upper = quantile(n_cond, .75, na.rm=TRUE)), keyby = year]


```




# Overall event numbers for indeterminate gender
```{r}

#Prevalence
first_year[, sum(end_bmm != 0), by = gender] #bmm
first_year[, sum(end_cmm != 0), by = gender] #cmm

#Incidence
first_year[, sum(end_bmm != 0 & cmm !=2), by = gender] #bmm
first_year[, sum(end_cmm != 0 & cmm !=2), by = gender] #cmm

#CF
first_year[,sum(end_bmm != 0 & censorreason ==1), by = gender] #bmm
first_year[,sum(end_cmm != 0 & censorreason ==1), by = gender] #cmm
```


# looking at censordates
```{r}
patient <- read_fst(data_dir_CPRD("patient.fst"), 
                    columns = c("patid", "pracid", "dob", "reg1yr"), 
                    as.data.table = TRUE) 
practice <- read_fst(data_dir_CPRD("practice.fst"), 
                    columns = c("pracid", "lcd"), 
                    as.data.table = TRUE) 
last_year[patient, on = 'patid', `:=` 
          (dob = i.dob, pracid = i.pracid, reg1yr = i.reg1yr)]
last_year[practice, on = 'pracid', lcd := i.lcd]
last_year[, enterstudy := reg1yr]
last_year[reg1yr < "2004-01-01", enterstudy := as.IDate("2004-01-01")]
last_year[enterstudy - dob < 6574, enterstudy := dob + 6574]
last_year[, daysinstudy := censordate - enterstudy]

#Censored for reason other than death #458,635
summary(last_year[censorreason == 0 & censordate < lcd & 
                    censordate < "2019-12-31" ])
#stayed until the end of the study/lcd of practice #462,413
summary(last_year[censorreason == 0 & 
                    (censordate == lcd |censordate >= "2019-12-31" ) ] )
#died #70,195 
summary(last_year[censorreason==1])



```

People who are censored (didnt die) are younger and have fewer conditions, more likely to have IMD as NA or be more deprived and more likely to be female. 