---
title: "Prevalence Graphs"
author: "Anna"
date: "01/04/2021"
output:   
  html_document:
    keep_md: true
---

```{r setup, include=FALSE}
library(ggplot2)
library(knitr)
library(fst)
library(tidyverse)
library(data.table)
if (!require("PHEindicatormethods")) install.packages("PHEindicatormethods")
library(PHEindicatormethods)
library(RColorBrewer)
library(ggthemes)
library(viridis)

#DSR package now archived
#packageurl <- "https://cran.r-project.org/src/contrib/Archive/dsr/dsr_0.2.2.tar.gz"
#install.packages(packageurl, repos=NULL, type="source")
library(dsr)
data.table::setDTthreads(10) #this is so that don't use all the processors 
```

Stuff for graphs 
```{r}
cbPalette <- c("#999999", "#E69F00", "#56B4E9", "#009E73", "#F0E442", "#0072B2", "#D55E00", "#CC79A7")

cbPaletteMod <- c("#999999", "#E69F00", "#56B4E9", "#009E73", "#D55E00", "#CC79A7")


mypalette <-brewer.pal(7,"GnBu")
mypalette #don't want the faintest 2 
mypalette <- c("#A8DDB5" ,"#7BCCC4" ,"#4EB3D3" ,"#2B8CBE", "#08589E")

mypalette2 <-brewer.pal(9,"GnBu")
mypalette2 #don't want the faintest 2 
mypalette2 <- c("#CCEBC5", "#A8DDB5" ,"#7BCCC4" ,"#4EB3D3" ,"#2B8CBE", "#0868AC" , "#084081")

xaxisyear <- c(2004, 2007, 2010, 2013, 2016, 2019)
```

# load the data
```{r echo=FALSE, out.width = '100%'}
data_dir_CPRD <-
  function(x = character(0))
    paste0("/mnt/", Sys.info()[["user"]],
      "/UoL/CPRD2019mm/Data May 2020/", x)

data_dir_lookup <-
  function(x = character(0))
    paste0("/mnt/", Sys.info()[["user"]],
           "/UoL/CPRD2019mm/Dictionaries etc/", x)


combi_mm <- read_fst(data_dir_CPRD("combi_mm_detailed.fst"), as.data.table = T)
combi_mm[, `:=` (
  imd = factor(imd, levels = c(1,2,3,4,5)), 
  bmm = factor(bmm),
  cmm = factor(cmm),
  cmm3plus = factor(cmm3plus))]

```


#Person time Prev
```{r}
#Making a summary table split by all the possible variables

prev_tab_pt <- rbind(combi_mm[,
                              .(
                                N = sum(bmm_days),
                                atrisk = sum(totaldays),
                                outcome = sum(bmm_days) / sum(totaldays) *
                                  100
                              ),
                              keyby = c("year",
                                        "gender",
                                        "agegrp5",
                                        "agegrp10_simple",
                                        "imd",
                                        "regionid",
                                        "region")][, type := "Basic"],
                     combi_mm[,
                              .(
                                N = sum(cmm_days),
                                atrisk = sum(totaldays),
                                outcome = sum(cmm_days) / sum(totaldays) *
                                  100
                              ),
                              keyby = c("year",
                                        "gender",
                                        "agegrp5",
                                        "agegrp10_simple",
                                        "imd",
                                        "regionid",
                                        "region")][, type := "Complex"],
                     combi_mm[,
                              .(
                                N = sum(cmm3plus_days),
                                atrisk = sum(totaldays),
                                outcome = sum(cmm3plus_days) / 
                                  sum(totaldays) * 100
                              ),
                              keyby = c("year",
                                        "gender",
                                        "agegrp5",
                                        "agegrp10_simple",
                                        "imd",
                                        "regionid",
                                        "region")][, type := "Complex 3+"])

```


#Overall crude prev by year
```{r}
prev_tab_pt[,
             sum(N)/sum(atrisk)*100, 
             keyby = c("year","type")]

prev_tab_pt[year %in% c(2004, 2019), sum(N)/sum(atrisk)*100, 
             keyby = c("year","type")]
```


#Overall crude prev by gender - table
```{r}
prev_tab_pt[,
             sum(N)/sum(atrisk)*100, 
             keyby = c("gender","type")]
```

# Crude prevalence by sex, MM type 
```{r}
ggplot(prev_tab_pt[gender != "Indeterminate" &
                     imd != "" & type != "Complex 3+",
                   sum(N) / sum(atrisk) * 100,
                   keyby = c("year", "type", "gender", "imd")],
       aes(year, V1, col = imd)) +
  geom_smooth(size = 0.5, se = FALSE) +
  facet_grid(type ~ gender) +
  expand_limits(y = 0) +
  scale_x_continuous(name = "Year",
                     limits = c(2003.5, 2020),
                     breaks = xaxisyear) +
  scale_y_continuous(name = "Crude Prevalence (%)") +
  theme_few() + theme(axis.text.x = element_text(size = 8),
                      legend.title = element_text(size = 10)) +
  scale_colour_manual(values = mypalette)  +
  labs(colour = "IMD")
#ggsave("prelim_graphs/prev_sex_imd.svg")

# With 3+ for SA - on side for easy comparison 
ggplot(prev_tab_pt[gender != "Indeterminate" & imd != "",
                   sum(N) / sum(atrisk) * 100,
                   keyby = c("year", "type", "gender", "imd")],
       aes(year, V1, col = imd)) +
  geom_smooth(size = 0.5, se = FALSE) +
  facet_grid(gender ~ type) +
  expand_limits(y = 0) +
  scale_x_continuous(name = "Year",
                     limits = c(2003.5, 2020),
                     breaks = xaxisyear) +
  scale_y_continuous(name = "Crude Prevalence (%)") +
  theme_few() + theme(axis.text.x = element_text(size = 8),
                      legend.title = element_text(size = 10)) +
  scale_colour_manual(values = mypalette)  +
  labs(colour = "IMD")
#ggsave("prelim_graphs/prelim_sa/prev_sex_imd_3p.svg") 

prev_tab_pt[gender != "Indeterminate" & imd != "",
            sum(N) / sum(atrisk) * 100,
            keyby = c("year", "type", "gender", "imd")]



```

# Crude prevalence by sex, agegrp, MM type 
```{r}
ggplot(prev_tab_pt[gender != "Indeterminate" & type != "Complex 3+",
                   (sum(N) / sum(atrisk) * 100),
                   keyby = c("year", "agegrp10_simple", "type", "gender")],
       aes(year, V1, col = agegrp10_simple)) +
  geom_smooth(size = 0.5, se = FALSE) +
  facet_grid(type ~ gender) +
  expand_limits(y = 0) +
  scale_x_continuous(name = "Year",
                     limits = c(2003.5, 2019.5),
                     breaks = xaxisyear) +
  scale_y_continuous(name = "Crude Prevalence",
                     labels = c("0%", "25%", "50%", "75%", "100%")) +
  theme_few() +
  theme(
    axis.text.x = element_text(size = 8),
    legend.background = element_rect(
      fill = '#FBFBFB',
      size = 0.5,
      linetype = "solid"
    ),
    legend.title = element_text(size = 10)
  ) +
  scale_colour_manual(values = mypalette2)  +
  labs(colour = "10-year\n age-group") +
  guides(colour = guide_legend(reverse = TRUE))
#ggsave("prelim_graphs/prev_sex_age.svg")

#With 3+ for SA - on side for readability
ggplot(prev_tab_pt[gender != "Indeterminate",
                   (sum(N) / sum(atrisk) * 100),
                   keyby = c("year", "agegrp10_simple", "type", "gender")],
       aes(year, V1, col = agegrp10_simple)) +
  geom_smooth(size = 0.5, se = FALSE) +
  facet_grid(gender ~ type) +
  expand_limits(y = 0) +
  scale_x_continuous(name = "Year",
                     limits = c(2003.5, 2019.5),
                     breaks = xaxisyear) +
  scale_y_continuous(name = "Crude Prevalence",
                     labels = c("0%", "25%", "50%", "75%", "100%")) +
  theme_few() +
  theme(
    axis.text.x = element_text(size = 8),
    legend.background = element_rect(
      fill = '#FBFBFB',
      size = 0.5,
      linetype = "solid"
    ),
    legend.title = element_text(size = 10)
  ) +
  scale_colour_manual(values = mypalette2)  +
  labs(colour = "10-year\n age-group")  +
  guides(colour = guide_legend(reverse = TRUE))
#ggsave("prelim_graphs/prelim_sa/prev_sex_age3p.svg")

#Looking at differences between men & women 
tmp <- prev_tab_pt[gender == "Men",
                   .(prev_m = sum(N) / sum(atrisk) * 100),
                   keyby = c("year", "type", "agegrp10_simple")]
tmp2 <- prev_tab_pt[gender == "Women",
                    .(prev_f = sum(N) / sum(atrisk) * 100),
                    keyby = c("year", "type", "agegrp10_simple")]
tmp[tmp2, on = c("year", "agegrp10_simple", "type"), prev_f := i.prev_f]
tmp[, fmdiff := prev_f - prev_m][order(fmdiff)]
```


# Heat map 

```{r}
#This is doing it on the crude rates, by  sex & imd & year.
require(tidyverse)
data <- prev_tab_pt[gender != "Indeterminate" & imd != "",
                    .(
                      N = sum(N),
                      atrisk = sum(atrisk),
                      outcome = sum(N) / sum(atrisk) * 100
                    ),
                    by = c("year", "imd", "gender", "agegrp10_simple", "type")][
                      , se := sqrt((outcome * (100 - outcome)) / atrisk)]

sii_sex_year_age_type <-
  as.data.table(
    phe_sii(
      group_by(data, year, gender, agegrp10_simple, type),
      imd,
      atrisk,
      value_type = 1,
      # for rates
      value = outcome,
      se = se,
      confidence = 0.95,
      rii = TRUE,
      type = "standard"
    )
  )


#As bar charts
ggplot(sii_sex_year_age_type[type != "Complex 3+"],
       aes(year, sii, fill = agegrp10_simple)) +
  geom_bar(stat = "identity", position = "dodge") +
  facet_grid(type ~ gender) +
  scale_x_continuous(name = "Year",
                     limits = c(2003.5, 2020),
                     breaks = xaxisyear) +
  scale_y_continuous(
    name = "Absolute Difference (Prevalence)",
    limits = c(-10, 20),
    labels = c("-10%", "0%", "10%", "20%")
  ) +
  theme_few() + theme(axis.text.x = element_text(size = 8),
                      legend.title = element_text(size = 10)) +
  scale_fill_brewer(type = 'seq',
                    direction = 1,
                    palette = 1) +    labs(fill = "10-year\n age-group") +
  guides(fill = guide_legend(reverse = TRUE))
#ggsave("prelim_graphs/prv_sii_ysat_bar1.svg")

#With 3+ for SA
ggplot(sii_sex_year_age_type,
       aes(year, sii, fill = agegrp10_simple)) +
  geom_bar(stat = "identity", position = "dodge") +
  facet_grid(gender ~ type) +
  scale_x_continuous(name = "Year",
                     limits = c(2003.5, 2020),
                     breaks = xaxisyear) +
  scale_y_continuous(
    name = "Absolute Difference (Prevalence)",
    limits = c(-10, 30),
    labels = c("-10%", "0%", "10%", "20%", "30%")
  ) +
  theme_few() + theme(axis.text.x = element_text(size = 8),
                      legend.title = element_text(size = 10)) +
  scale_fill_brewer(type = 'seq',
                    direction = 1,
                    palette = 1) +    
  labs(fill = "10-year\n age-group") +
  guides(colour = guide_legend(reverse = TRUE))
#ggsave("prelim_graphs/prelim_sa/prv_sii_ysat_bar1.svg") 


```


# Age-standardised

```{r}
eustandpop <- fread(file = data_dir_lookup("eustandardpop5yrband.csv"), 
                    header = TRUE, sep = ",") #European Standardised pop 
eustandpop[, gender := as.factor(ifelse(gender== "Male", "Men", "Women"))] 
setnames(eustandpop, c("stand.pop", "age.group"), c("pop", "agegrp5"))


dsr_a_imd <- function(x, y){
  DT <- prev_tab_pt[imd == x & gender != "Indeterminate" & type == y, 
                                .(N = sum(N), atrisk = sum(atrisk)),
                    by = c("year", "gender", "agegrp5")]
  z <-  rbind(
  as.data.table( dsr(
    data= DT[gender == "Men", ],
    event=N,
    fu=atrisk,
    subgroup=year, age.group, type,
    refdata=eustandpop[gender == "Men",],
    method="gamma",
    sig=0.95,
    mp=100,
    decimals=4))[, gender := "Men"], 
  as.data.table( dsr(
        data= DT[gender == "Women", ],
        event=N,
        fu=atrisk,
        subgroup=year, age.group, type,
        refdata=eustandpop[gender == "Women",],
        method="gamma",
        sig=0.95,
        mp=100,
        decimals=4))[, gender := "Women"] )
  z[,`:=` (imd = x)]
  z
}

bmm_prv_ast_imd <- rbind(
  dsr_a_imd(1, "Basic"),
  dsr_a_imd(2, "Basic"),
  dsr_a_imd(3, "Basic"),
  dsr_a_imd(4, "Basic"),
  dsr_a_imd(5, "Basic")
)

bmm_prv_ast_imd[, imd := factor(imd)]
setnames(
  bmm_prv_ast_imd,
  c("Subgroup", "Std Rate (per 100)", "95% LCL (Std)", "95% UCL (Std)"),
  c("year", "standoutcome", "standlower", "standupper")
)

cmm_prv_ast_imd <- rbind(
  dsr_a_imd(1, "Complex"),
  dsr_a_imd(2, "Complex"),
  dsr_a_imd(3, "Complex"),
  dsr_a_imd(4, "Complex"),
  dsr_a_imd(5, "Complex")
)

cmm_prv_ast_imd[, imd := factor(imd)]
setnames(
  cmm_prv_ast_imd,
  c("Subgroup", "Std Rate (per 100)", "95% LCL (Std)", "95% UCL (Std)"),
  c("year", "standoutcome", "standlower", "standupper")
)

cmm3plus_prv_ast_imd <- rbind(
  dsr_a_imd(1, "Complex 3+"),
  dsr_a_imd(2, "Complex 3+"),
  dsr_a_imd(3, "Complex 3+"),
  dsr_a_imd(4, "Complex 3+"),
  dsr_a_imd(5, "Complex 3+")
)

cmm3plus_prv_ast_imd[, imd := factor(imd)]
setnames(
  cmm3plus_prv_ast_imd,
  c("Subgroup", "Std Rate (per 100)", "95% LCL (Std)", "95% UCL (Std)"),
  c("year", "standoutcome", "standlower", "standupper")
)

rm(dsr_a_imd)

combi_prv_ast_imd <-
  rbind(bmm_prv_ast_imd[, type := "Basic"], 
        cmm_prv_ast_imd[, type := "Complex"], 
        cmm3plus_prv_ast_imd[, type := "Complex 3+"])

ggplot(combi_prv_ast_imd[type != "Complex 3+"]) + theme_few() +
  geom_smooth(
    mapping = aes(x = year, y = standoutcome, colour = imd),
    se = FALSE,
    size = 0.5
  ) +
  facet_grid(type ~ gender) +
  expand_limits(y = 0) +
  scale_x_continuous(name = "Year",
                     limits = c(2003.5, 2019.5),
                     breaks = xaxisyear) +
  scale_y_continuous(name = "Age-standardised prevalence",
                     labels = c("0%", "20%", "40%", "60%")) +
  theme(
    axis.text.x = element_text(size = 8),
    legend.background = element_rect(
      fill = '#FBFBFB',
      size = 0.5,
      linetype = "solid"
    ),
    legend.title = element_text(size = 10)
  ) +
  scale_colour_manual(values = mypalette)  +
  labs(colour = "IMD")  +
  guides(colour = guide_legend(reverse = TRUE))

#ggsave("prelim_graphs/prv_combi_asst_sex_imd.svg")

# with 3+ for SA
ggplot(combi_prv_ast_imd) + theme_few() +
  geom_smooth(
    mapping = aes(x = year, y = standoutcome, colour = imd),
    se = FALSE,
    size = 0.5
  ) +
  facet_grid(gender ~ type) +
  expand_limits(y = 0) +
  scale_x_continuous(name = "Year",
                     limits = c(2003.5, 2019.5),
                     breaks = xaxisyear) +
  scale_y_continuous(name = "Age-standardised prevalence",
                     labels = c("0%", "20%", "40%", "60%")) +
  theme(
    axis.text.x = element_text(size = 8),
    legend.background = element_rect(
      fill = '#FBFBFB',
      size = 0.5,
      linetype = "solid"
    ),
    legend.title = element_text(size = 10)
  ) +
  scale_colour_manual(values = mypalette)  +
  labs(colour = "IMD")  +
  guides(colour = guide_legend(reverse = TRUE))
#ggsave("prelim_graphs/prelim_sa/prv_combi_asst_sex_imd.svg")

combi_prv_ast_imd

```


# Age-sex standardised by year by imd

```{r}


dsr_as_imd <- function(x, y) {
  DT <- prev_tab_pt[imd == x & gender != "Indeterminate" & type == y,
                    .(N = sum(N), atrisk = sum(atrisk)),
                    by = c("year", "gender", "agegrp5")]
  z <- as.data.table(
    dsr(
      data = DT,
      event = N,
      fu = atrisk,
      subgroup = year,
      age.group,
      type,
      refdata = eustandpop,
      method = "gamma",
      sig = 0.95,
      mp = 100,
      decimals = 4
    )
  )
  z[, `:=` (imd = x)]
  z
}
bmm_prv_asst_imd <- rbind(
  dsr_as_imd(1, "Basic"),
  dsr_as_imd(2, "Basic"),
  dsr_as_imd(3, "Basic"),
  dsr_as_imd(4, "Basic"),
  dsr_as_imd(5, "Basic")
)

bmm_prv_asst_imd[, imd := factor(imd)]
setnames(
  bmm_prv_asst_imd,
  c("Subgroup", "Std Rate (per 100)", "95% LCL (Std)", "95% UCL (Std)"),
  c("year", "standoutcome", "standlower", "standupper")
)

cmm_prv_asst_imd <- rbind(
  dsr_as_imd(1, "Complex"),
  dsr_as_imd(2, "Complex"),
  dsr_as_imd(3, "Complex"),
  dsr_as_imd(4, "Complex"),
  dsr_as_imd(5, "Complex")
)

cmm_prv_asst_imd[, imd := factor(imd)]
setnames(
  cmm_prv_asst_imd,
  c("Subgroup", "Std Rate (per 100)", "95% LCL (Std)", "95% UCL (Std)"),
  c("year", "standoutcome", "standlower", "standupper")
)

cmm3plus_prv_asst_imd <- rbind(
  dsr_as_imd(1, "Complex 3+"),
  dsr_as_imd(2, "Complex 3+"),
  dsr_as_imd(3, "Complex 3+"),
  dsr_as_imd(4, "Complex 3+"),
  dsr_as_imd(5, "Complex 3+")
)

cmm3plus_prv_asst_imd[, imd := factor(imd)]
setnames(
  cmm3plus_prv_asst_imd,
  c("Subgroup", "Std Rate (per 100)", "95% LCL (Std)", "95% UCL (Std)"),
  c("year", "standoutcome", "standlower", "standupper")
)
rm(dsr_as_imd)


combi_prv_asst_imd <-
  rbind(bmm_prv_asst_imd[, type := "Basic"], 
        cmm_prv_asst_imd[, type := "Complex"], 
        cmm3plus_prv_asst_imd[, type := "Complex 3+"])

ggplot(combi_prv_asst_imd) + theme_few() +
  geom_smooth(
    mapping = aes(x = year, y = standoutcome, colour = imd),
    se = FALSE,
    size = 0.5
  ) +
  facet_grid(cols = vars(type)) +
  expand_limits(y = 0) +
  scale_x_continuous(name = "Year",
                     limits = c(2003.5, 2019.5),
                     breaks = xaxisyear) +
  scale_y_continuous(
    name = "Age-standardised prevalence",
    limits = c(0, 60),
    labels = c("0%", "20%", "40%", "60%")
  ) +
  theme(
    axis.text.x = element_text(size = 8),
    legend.background = element_rect(
      fill = '#FBFBFB',
      size = 0.5,
      linetype = "solid"
    ),
    legend.title = element_text(size = 10)
  ) +
  scale_colour_manual(values = mypalette)  +
  labs(colour = "IMD")

#ggsave("prelim_graphs/prv_combi_asst_sex_imd.png")

combi_prv_asst_imd
```

# Age-sex standardised by year by region (this is for the map)

```{r}


dsr_as_reg <- function(x, y) {
  DT <-
    prev_tab_pt[regionid == x & gender != "Indeterminate" & type == y,
                .(N = sum(N), atrisk = sum(atrisk)),
                by = c("year", "gender", "agegrp5")]
  z <- as.data.table(
    dsr(
      data = DT,
      event = N,
      fu = atrisk,
      subgroup = year,
      age.group,
      type,
      refdata = eustandpop,
      method = "gamma",
      sig = 0.95,
      mp = 100,
      decimals = 4
    )
  )
  z[, `:=` (regionid = x)]
  z
}

bmm_prv_asst_reg <- rbind(
  dsr_as_reg(1, "Basic"),
  dsr_as_reg(2, "Basic"),
  dsr_as_reg(3, "Basic"),
  dsr_as_reg(4, "Basic"),
  dsr_as_reg(5, "Basic"),
  dsr_as_reg(6, "Basic"),
  dsr_as_reg(7, "Basic"),
  dsr_as_reg(8, "Basic"),
  dsr_as_reg(9, "Basic"),
  dsr_as_reg(10, "Basic")
)

setnames(
  bmm_prv_asst_reg,
  c("Subgroup", "Std Rate (per 100)", "95% LCL (Std)", "95% UCL (Std)"),
  c("year", "standoutcome", "standlower", "standupper")
)

cmm_prv_asst_reg <- rbind(
  dsr_as_reg(1, "Complex"),
  dsr_as_reg(2, "Complex"),
  dsr_as_reg(3, "Complex"),
  dsr_as_reg(4, "Complex"),
  dsr_as_reg(5, "Complex"),
  dsr_as_reg(6, "Complex"),
  dsr_as_reg(7, "Complex"),
  dsr_as_reg(8, "Complex"),
  dsr_as_reg(9, "Complex"),
  dsr_as_reg(10, "Complex")
)

setnames(
  cmm_prv_asst_reg,
  c("Subgroup", "Std Rate (per 100)", "95% LCL (Std)", "95% UCL (Std)"),
  c("year", "standoutcome", "standlower", "standupper")
)

cmm3plus_prv_asst_reg <- rbind(
  dsr_as_reg(1, "Complex 3+"),
  dsr_as_reg(2, "Complex 3+"),
  dsr_as_reg(3, "Complex 3+"),
  dsr_as_reg(4, "Complex 3+"),
  dsr_as_reg(5, "Complex 3+"),
  dsr_as_reg(6, "Complex 3+"),
  dsr_as_reg(7, "Complex 3+"),
  dsr_as_reg(8, "Complex 3+"),
  dsr_as_reg(9, "Complex 3+"),
  dsr_as_reg(10, "Complex 3+")
)

setnames(
  cmm3plus_prv_asst_reg,
  c("Subgroup", "Std Rate (per 100)", "95% LCL (Std)", "95% UCL (Std)"),
  c("year", "standoutcome", "standlower", "standupper")
)
rm(dsr_as_reg)

combi_prv_asst_reg <-
  rbind(bmm_prv_asst_reg[, type := "Basic"], 
        cmm_prv_asst_reg[, type := "Complex"], 
        cmm3plus_prv_asst_reg[, type := "Complex 3+"])

data_dir_lookup <-
  function(x = character(0))
    paste0("/mnt/", Sys.info()[["user"]],
           "/UoL/CPRD2019mm/Dictionaries etc/", x)

cprd_reg <- fread(file = data_dir_lookup("Region.txt"),
                  header = TRUE,
                  sep = "\t")

combi_prv_asst_reg[, regionid := as.integer(regionid)]
combi_prv_asst_reg[cprd_reg, on = 'regionid', region := i.Description]
combi_prv_asst_reg[, region := factor(
  region,
  levels = c(
    "London",
    "South West",
    "South Central",
    "South East Coast",
    "West Midlands",
    "East Midlands",
    "East of England",
    "North West",
    "Yorkshire And The Humber",
    "North East"
  ),
  labels = c(
    "London",
    "South West",
    "South Central",
    "South East Coast",
    "West Midlands",
    "East Midlands",
    "East of England",
    "North West",
    "Yorkshire And The Humber",
    "North East"
  )
)]
rm(cprd_reg)


combi_prv_asst_reg
```

# Age-sex standardised by imd

```{r}


dsr_ass_imd_ny <- function(x){
  DT <- prev_tab_pt[imd != "" & gender != "Indeterminate" & type == x, 
                                .(N = sum(N), atrisk = sum(atrisk)),
                                                                by = c("imd", "gender", "agegrp5")]
  z <- as.data.table( dsr(
    data= DT,
    event=N,
    fu=atrisk,
    subgroup=imd, age.group, type,
    refdata=eustandpop,
    method="gamma",
    sig=0.95,
    mp=100,
    decimals=4))
  z[,`:=` (type = x)]
  z
}

prv_asst_imd_ny <- rbind(
  dsr_ass_imd_ny("Basic"), dsr_ass_imd_ny("Complex"), dsr_ass_imd_ny("Complex 3+"))

setnames(prv_asst_imd_ny, 
         c("Subgroup", "Std Rate (per 100)", "95% LCL (Std)", "95% UCL (Std)"), 
         c("imd", "standoutcome", "standlower", "standupper"))
prv_asst_imd_ny[, imd := factor(imd)]
prv_asst_imd_ny[, .(imd, type, `Crude Rate (per 100)`, standoutcome)]
```

# Overall age-sex standardised by year
```{r}

dsr_assy <- function(x) {
  DT <-
    prev_tab_pt[imd != "" & gender != "Indeterminate" & year == x,
                .(N = sum(N), atrisk = sum(atrisk)),
                by = c("type", "gender", "agegrp5")]
  z <- as.data.table(
    dsr(
      data = DT,
      event = N,
      fu = atrisk,
      subgroup = type,
      refdata = eustandpop,
      method = "gamma",
      sig = 0.95,
      mp = 100,
      decimals = 4
    )
  )
  z[, `:=` (year = as.numeric(x))]
  z
}


 prv_assy <- data.table()
  for(i in (2004:2019)){ 
    subtab <- dsr_assy(i)
    prv_assy <- rbind(prv_assy, subtab)
    rm(subtab)
  }
  rm(i)
  

setnames(prv_assy, 
         c("Subgroup", "Std Rate (per 100)", "95% LCL (Std)", "95% UCL (Std)"), 
         c("type", "standoutcome", "standlower", "standupper"))
prv_assy
prv_assy[year %in% c(2004, 2019)]
```


# Indeterminate gender crude prev
```{r}
prev_tab_pt[,
             sum(N)/sum(atrisk)*100, 
             keyby = c("type", "gender")]
```


# Regional graphs - crude

```{r}
#Download some important packages
library(maps)
if (!require("mapdata")) install.packages("mapdata")
library(mapdata)
library(maptools)
library(rgdal)
library(ggmap)
library(ggplot2)
library(rgeos)
library(broom)
library(plyr)
library(data.table)
library(cowplot)



#Load the shapefile - make sure you change the filepath to where you saved the shapefiles
shapefile <- readOGR(dsn= data_dir_lookup("SHA boundaries"), 
                     layer="SHA_DEC_2011_EN_BFC")

#Downloaded from here https://datashare.is.ed.ac.uk/handle/10283/2559?show=full 
#This is by strategic health authority

#Reshape for ggplot2 using the Broom package
mapdata <- tidy(shapefile, region="SHA11NM") #This might take a few minutes

#Check the shapefile has loaded correctly by plotting an outline map of the UK
gg <-
  ggplot() + geom_polygon(
    data = mapdata,
    aes(x = long, y = lat, group = group),
    color = "#FFFFFF",
    size = 0.25
  )
gg <-
  gg + coord_fixed(1) #This gives the map a 1:1 aspect ratio to prevent 
#the map from appearing squashed
print(gg)

regional_prev <- prev_tab_pt[!is.na(region), 
         .(outcome = sum(N)/sum(atrisk)*100), 
         keyby = c("year", "region", "regionid", "type")]

cprd_reg <- fread(file = data_dir_lookup("Region.txt"), 
                              header = TRUE, sep = "\t") 
cprd_reg
map_reg <- fread(file = data_dir_lookup(
  "Map_Region__December_2015__Boundaries.csv"), 
                 header = TRUE, sep = ",") 
map_reg

#Matching the cprd_reg names to the map_reg
cprd_reg[Description == "Yorkshire And The Humber", 
         Description := "Yorkshire and the Humber"]
#checking they match
cprd_reg[Description %in% shapefile$SHA11NM]
regional_prev[, regionid := as.numeric(regionid)]
regional_prev[cprd_reg, on ='regionid', id := i.Description]



groupprvfn <- function(x) { 
  lev <- c("10-14%", "15-19%",
           "20-24%", "25-29%",
           "30-34%", "35-39%",
           "40-44%", "45-49%",
           "50-54%", "55-59%")
  x[, Prevalence := fcase(
    between(outcome, 54.99, 59.99), factor(lev[10], levels = lev),
    between(outcome, 49.99, 54.99), factor(lev[9], levels = lev),
    between(outcome, 44.99, 49.99), factor(lev[8], levels = lev),
    between(outcome, 39.99, 44.99), factor(lev[7], levels = lev),
    between(outcome, 34.99, 39.99), factor(lev[6], levels = lev),
    between(outcome, 29.99, 34.99), factor(lev[5], levels = lev),
    between(outcome, 24.99, 29.99), factor(lev[4], levels = lev),
    between(outcome, 19.99, 24.99), factor(lev[3], levels = lev),
    between(outcome, 14.99, 19.99), factor(lev[2], levels = lev),
    between(outcome, 9.99, 14.99), factor(lev[1], levels = lev)
  )
  ]
}
groupprvfn(regional_prev)

data <- regional_prev[year %in% c(2004, 2019), #subsetting for merging
                           .(year, type, id, outcome, Prevalence)]

df <- join(mapdata, data, by = "id")
df <- as.data.table(df)


#Only graphing complex MM for 

cats <- c("10-14%", "15-19%",
           "20-24%", "25-29%",
           "30-34%", "35-39%")


gg_cat <- ggplot() + 
  geom_polygon(data = df[year == 2004 & type == "Complex"], 
               aes(x = long, y = lat, group = group, fill = Prevalence), 
               color = "#FFFFFF", size = 0.25) 

gg_cat <- gg_cat + 
  scale_fill_brewer(type = "seq", direction = 1,  limits = c(cats)) +
  coord_fixed(1) + 
  theme_minimal()+ 
  theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank())+ 
  theme(axis.title.x=element_blank(), axis.text.x = element_blank(), axis.ticks.x = element_blank())+ 
  theme(axis.title.y=element_blank(), axis.text.y = element_blank(), axis.ticks.y = element_blank())

gg2_cat <- ggplot() + 
  geom_polygon(data = df[year == 2019 & type == "Complex"], 
               aes(x = long, y = lat, group = group, fill = Prevalence), 
               color = "#FFFFFF", size = 0.25)

gg2_cat <- gg2_cat +
  scale_fill_brewer(type = "seq",
                    direction = 1,
                    limits = c(cats)) +
  coord_fixed(1) +
  theme_minimal() +
  theme(panel.grid.major = element_blank(),
        panel.grid.minor = element_blank()) +
  theme(
    axis.title.x = element_blank(),
    axis.text.x = element_blank(),
    axis.ticks.x = element_blank()
  ) +
  theme(
    axis.title.y = element_blank(),
    axis.text.y = element_blank(),
    axis.ticks.y = element_blank()
  )



plot_grid(gg_cat, gg2_cat, labels = c('2004', '2019'))

#ggsave("prelim_graphs/prv_cmm_regmap.png") 


```


# Regional graphs age sex standardised
```{r}
combi_prv_asst_reg [cprd_reg, on = 'regionid', id := i.Description]
setnames(combi_prv_asst_reg, "standoutcome", "outcome")
groupprvfn(combi_prv_asst_reg)
data <-
  combi_prv_asst_reg[year %in% c(2004, 2019), #subsetting for merging
                     .(year, type, id, outcome, Prevalence)]
df2 <- join(mapdata, data, by = "id")
df2 <- as.data.table(df2)

cats2 <- c("10-14%", "15-19%",
           "20-24%", "25-29%",
           "30-34%", "35-39%")


#with only 1 legend
df2_c <- df2[type == "Complex"]

gg3 <- ggplot() +
  geom_polygon(
    data = df2_c,
    aes(
      x = long,
      y = lat,
      group = group,
      fill = Prevalence
    ),
    color = "#FFFFFF",
    size = 0.25
  ) +
  facet_grid(cols = vars(year))

gg3 <- gg3 +
  scale_fill_brewer(type = "seq",
                    direction = 1,
                    limits = c(cats2)) +
  coord_fixed(1) +
  theme_minimal() +
  theme(panel.grid.major = element_blank(),
        panel.grid.minor = element_blank()) +
  theme(
    axis.title.x = element_blank(),
    axis.text.x = element_blank(),
    axis.ticks.x = element_blank()
  ) +
  theme(
    axis.title.y = element_blank(),
    axis.text.y = element_blank(),
    axis.ticks.y = element_blank()
  )
gg3
#ggsave("prelim_graphs/prv_cmm_regmap_stand.svg") 
```

# Basic - for appendx
```{r}
df2_b <- df2[type == "Basic"]

gg4 <- ggplot() +
  geom_polygon(
    data = df2_b,
    aes(
      x = long,
      y = lat,
      group = group,
      fill = Prevalence
    ),
    color = "#FFFFFF",
    size = 0.25
  ) +
  facet_grid(cols = vars(year))

gg4 <- gg4 +
  scale_fill_brewer(type = "seq", direction = 1) +
  coord_fixed(1) +
  theme_minimal() +
  theme(panel.grid.major = element_blank(),
        panel.grid.minor = element_blank()) +
  theme(
    axis.title.x = element_blank(),
    axis.text.x = element_blank(),
    axis.ticks.x = element_blank()
  ) +
  theme(
    axis.title.y = element_blank(),
    axis.text.y = element_blank(),
    axis.ticks.y = element_blank()
  )
gg4
#ggsave("prelim_graphs/prelim_sa/prv_bmm_regmap_stand.svg") 
```


# Complex 3+ - for appendx
```{r}
df2_c3p <- df2[type == "Complex 3+"]

gg5 <- ggplot() + 
    geom_polygon(data = df2_c3p, 
                 aes(x = long, y = lat, group = group, fill = Prevalence), 
                 color = "#FFFFFF", size = 0.25) +
facet_grid(cols = vars(year)) 

gg5 <- gg5 +
  scale_fill_brewer(type = "seq", direction = 1) +
  coord_fixed(1) +
  theme_minimal() +
  theme(panel.grid.major = element_blank(),
        panel.grid.minor = element_blank()) +
  theme(
    axis.title.x = element_blank(),
    axis.text.x = element_blank(),
    axis.ticks.x = element_blank()
  ) +
  theme(
    axis.title.y = element_blank(),
    axis.text.y = element_blank(),
    axis.ticks.y = element_blank()
  )
gg5
#ggsave("prelim_graphs/prelim_sa/prv_cmm3p_regmap_stand.svg") 
```



# Point prevalence estimates
```{r}

#Making a summary table split by all the possible variables

prev_tab_point <- rbind(combi_mm[,
                                 .(
                                   N = sum(pointprev_bmm),
                                   atrisk = sum(pointprev_dnom),
                                   outcome = sum(pointprev_bmm) / 
                                     sum(pointprev_dnom) * 100
                                 ),
                                 keyby = c("year",
                                           "gender",
                                           "agegrp5",
                                           "agegrp10_simple",
                                           "imd",
                                           "regionid",
                                           "region")][, type := "Basic"],
                        combi_mm[,
                                 .(
                                   N = sum(pointprev_cmm),
                                   atrisk = sum(pointprev_dnom),
                                   outcome = sum(pointprev_cmm) / 
                                     sum(pointprev_dnom) * 100
                                 ),
                                 keyby = c("year",
                                           "gender",
                                           "agegrp5",
                                           "agegrp10_simple",
                                           "imd",
                                           "regionid",
                                           "region")][, type := "Complex"])

ggplot(prev_tab_point[gender != "Indeterminate",
                      (sum(N) / sum(atrisk) * 100),
                      keyby = c("year", "agegrp10_simple", "gender", "type")],
       aes(year, V1, col = agegrp10_simple)) +
  geom_smooth(size = 0.5, se = FALSE) +
  facet_grid(type ~ gender) +
  expand_limits(y = 0) +
  scale_x_continuous(name = "Year",
                     limits = c(2003.5, 2019.5),
                     breaks = xaxisyear) +
  scale_y_continuous(name = "Crude Prevalence",
                     labels = c("0%", "25%", "50%", "75%", "100%")) +
  theme_few() +
  theme(
    axis.text.x = element_text(size = 8),
    legend.background = element_rect(
      fill = '#FBFBFB',
      size = 0.5,
      linetype = "solid"
    ),
    legend.title = element_text(size = 10)
  ) +
  scale_colour_manual(values = mypalette2)  +
  labs(colour = "10-year\n age-group") #+
#ggsave("prelim_graphs/prev_sex_age.svg") 



dsr_a_imd <- function(x, y) {
  DT <-
    prev_tab_point[imd == x & gender != "Indeterminate" & type == y,
                   .(N = sum(N), atrisk = sum(atrisk)),
                   by = c("year", "gender", "agegrp5")]
  z <-  rbind(as.data.table(
    dsr(
      data = DT[gender == "Men",],
      event = N,
      fu = atrisk,
      subgroup = year,
      age.group,
      type,
      refdata = eustandpop[gender == "Men", ],
      method = "gamma",
      sig = 0.95,
      mp = 100,
      decimals = 4
    )
  )[, gender := "Men"],
  as.data.table(
    dsr(
      data = DT[gender == "Women",],
      event = N,
      fu = atrisk,
      subgroup = year,
      age.group,
      type,
      refdata = eustandpop[gender == "Women", ],
      method = "gamma",
      sig = 0.95,
      mp = 100,
      decimals = 4
    )
  )[, gender := "Women"])
  z[, `:=` (imd = x)]
  z
}

bmm_pointprv_ast_imd <- rbind(
  dsr_a_imd(1, "Basic"),
  dsr_a_imd(2, "Basic"),
  dsr_a_imd(3, "Basic"),
  dsr_a_imd(4, "Basic"),
  dsr_a_imd(5, "Basic")
)


cmm_pointprv_ast_imd <- rbind(
  dsr_a_imd(1, "Complex"),
  dsr_a_imd(2, "Complex"),
  dsr_a_imd(3, "Complex"),
  dsr_a_imd(4, "Complex"),
  dsr_a_imd(5, "Complex")
)


combi_pointprv_ast_imd <-
  rbind(bmm_pointprv_ast_imd[, type := "Basic"],
        cmm_pointprv_ast_imd[, type := "Complex"])

combi_pointprv_ast_imd[, imd := factor(imd)]
setnames(
  combi_pointprv_ast_imd,
  c("Subgroup", "Std Rate (per 100)", "95% LCL (Std)", "95% UCL (Std)"),
  c("year", "standoutcome", "standlower", "standupper")
)


ggplot(combi_pointprv_ast_imd) + theme_few() +
  geom_smooth(
    mapping = aes(x = year, y = standoutcome, colour = imd),
    se = FALSE,
    size = 0.5
  ) +
  facet_grid(type ~ gender) +
  expand_limits(y = 0) +
  scale_x_continuous(name = "Year",
                     limits = c(2003.5, 2019.5),
                     breaks = xaxisyear) +
  scale_y_continuous(name = "Age-standardised point prevalence",
                     labels = c("0%", "20%", "40%", "60%")) +
  theme(
    axis.text.x = element_text(size = 8),
    legend.background = element_rect(
      fill = '#FBFBFB',
      size = 0.5,
      linetype = "solid"
    ),
    legend.title = element_text(size = 10)
  ) +
  scale_colour_manual(values = mypalette)  + labs(colour = "IMD")
#ggsave("prelim_graphs/prelim_sa/pointprv_agest.svg")



# Overall age-sex standardised by year

dsr_assy <- function(x) {
  DT <-
    prev_tab_point[imd != "" & gender != "Indeterminate" & year == x,
                   .(N = sum(N), atrisk = sum(atrisk)),
                   by = c("type", "gender", "agegrp5")]
  z <- as.data.table(
    dsr(
      data = DT,
      event = N,
      fu = atrisk,
      subgroup = type,
      refdata = eustandpop,
      method = "gamma",
      sig = 0.95,
      mp = 100,
      decimals = 4
    )
  )
  z[, `:=` (year = as.numeric(x))]
  z
}


pointprv_assy <- data.table()
for (i in (2004:2019)) {
  subtab <- dsr_assy(i)
  pointprv_assy <- rbind(pointprv_assy, subtab)
  rm(subtab)
}
rm(i)


setnames(
  pointprv_assy,
  c("Subgroup", "Std Rate (per 100)", "95% LCL (Std)", "95% UCL (Std)"),
  c("type", "standoutcome", "standlower", "standupper")
)
pointprv_assy
pointprv_assy[year %in% c(2004, 2019)]
```

